\documentclass[prl,12pt,onecolumn,nofootinbib,notitlepage,english,superscriptaddress]{revtex4-1}
\renewcommand{\rmdefault}{cmr}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{color}
\usepackage{babel}
\usepackage{latexsym}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{times}   %% Times Roman font
\usepackage{esint}
\usepackage{subfigure}
\usepackage{verbatim}
\usepackage{braket}
\usepackage{footmisc}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=false,colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=red]{hyperref}
\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\special{papersize=\the\paperwidth,\the\paperheight}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\@ifundefined{textcolor}{}
{%
 \definecolor{BLACK}{gray}{0}
 \definecolor{WHITE}{gray}{1}
 \definecolor{RED}{rgb}{1,0,0}
 \definecolor{GREEN}{rgb}{0,1,0}
 \definecolor{BLUE}{rgb}{0,0,1}
 \definecolor{CYAN}{cmyk}{1,0,0,0}
 \definecolor{MAGENTA}{cmyk}{0,1,0,0}
 \definecolor{YELLOW}{cmyk}{0,0,1,0}
}

\@ifundefined{date}{}{\date{}}
\AtBeginDocument{
  \def\labelitemi{\(\rhd\)}
}
\makeatother

\setlength{\belowcaptionskip}{-7pt}
\newcommand{\SAVE}[1]{}
\newcommand{\HJC}[1]{{\color{RED}{\bf HJC: #1}}}
\newcommand{\prlsec}[1]{\emph{#1---}}
\newcommand{\Ncal}{{\mathcal N}}
\newcommand{\T}{{\mathbf{T}}}
\newcommand{\Jbq}{{J_{bq}}}
\newcommand{\DK}[1]{{\color{BLUE}{\bf DK: #1}}}

\begin{document}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand\abstractname{}
\title{From real materials to model lattice Hamiltonians: multi-scale modelling of strongly correlated electronic systems 
       with information from many body wavefunctions}

\author{Hitesh J. Changlani}
\affiliation{Department of Physics and Astronomy, Johns Hopkins University, Baltimore, Maryland 21218, USA}
\author{Huihuo Zheng}
\affiliation{
Argonne Leadership Computing Facility, Argonne National Laboratory, 9700 South Cass Avenue, Lemont, 60439, Illinois, USA}
\author{Kiel Williams}
\affiliation{Department of Physics and Institute for Condensed Matter Theory, University of Illinois at Urbana-Champaign, 
1110 West Green St, Urbana IL 61801, USA}
\author{Brian Busemeyer}
\affiliation{Department of Physics and Institute for Condensed Matter Theory, University of Illinois at Urbana-Champaign, 
1110 West Green St, Urbana IL 61801, USA}
\author{Lucas K. Wagner}
\affiliation{Department of Physics and Institute for Condensed Matter Theory, University of Illinois at Urbana-Champaign, 
1110 West Green St, Urbana IL 61801, USA}
\date{\today}
\maketitle

\textbf{
Given a realistic material with all its intrinsic complications, how does one develop a 
simple reliable model for understanding its properties? Theoretical insight has been the key driver of 
this process leading to simple few-band pictures. When the interactions are comparable 
to or much stronger than the kinetic energy, it is convenient to adopt the real space lattice 
approach and work with Hubbard or Kanamori type Hamiltonians involving only the low energy electrons. 
But this is no easy task, since the effective Hamiltonian involves a considerable renormalization of parameters with respect 
to the bare Coulomb values. While the kinetic energy is dominated by contributions from bands or states energetically 
near the Fermi level, screened interactions depend on states even \emph{far} away from it, leading to Hubbard U's 
that have been traditionally hard to reliably determine. 
Here we discuss an approach that treats the kinetic and potential parts of the Hamiltonian 
democratically and one that provides a transparent way of obtaining effective Hamiltonians using data from many body wavefunctions, 
and whose validity can be systematically checked.
We emphasize that determining the effective Hamiltonian reliably is crucial for 
several applications in physics and chemistry not only for quantitative accuracy, but also a 
correct qualitative picture of strongly correlated materials. 
}

\input{intro.tex}
%\input{theory.tex}
\section{Theory: Compression of the Energy Functional/Criteria for Hamiltonian matching}
\HJC{Lucas handling this section......Look at theory.tex. Could use commented out snippets explaining Ax=E etc and criteria}
\input{threeband.tex}
\input{hchain.tex}
\input{honeycomb.tex}

\section{Conclusion and Future prospects}
To summarize, we have explained the AIDMD method that we have been developing, primarily in conjunction with the 
\emph{ab-initio} QMC approach. The practical motivation is to take data from first principles (continuum) 
calculations as inputs for lattice model (discrete) methods. 
Since one is mapping a many body problem to a few body one, the relevant quantities of interest are 
the reduced density matrices associated with the many-body wavefunctions projected to a one-body space. 
The density matrix based approach is appealing as it is democratic 
in the determination of the hopping and interaction parts of the effective Hamiltonian, and provides multiple 
checks on its validity. We have discussed three representative examples to discuss the conceptual and algorithmic aspects of AIDMD. 

What sort of accuracy should one expect with effective Hamiltonians and AIDMD? The holy grail of quantum chemistry 
and electronic structure is to obtain an energy of 1 mHa (0.027 eV) per atom; this 
remains an open challenge despite decades of work. The effective Hamiltonian approach largely ameliorates this problem,  only 
relative energies are important, neither the total energy (nor its accuracy) is of particularly 
fundamental interest. This, of course, is contingent on the cancellation of the large energy associated with the core electrons, 
whose role is to primarily renormalize the interactions between the active (valence) electrons. This aspect can also 
be a major limitation when the core and active electrons are strongly entangled i.e when the separation of energy 
scales is not prominent (in which case an energy-dependent description may be needed). 
Thus, it is possibly best suited for extended systems (solids) where this separation exists. 
In our view, AIDMD (and downfolding in general) opens up many avenues for determination of physical 
quantities not easily calculated in ground state approaches. Our experience suggests excitation spectra 
can be accurately determined to 0.2 eV (or less)~\cite{Changlani2015} and potentially improved with 
more \emph{ab-initio} data and more refined effective Hamiltonians.  
%It is also possibly best suited for extended systems (solids) where there is often a clear separation of energy scales associated with the 
%core, active (valence) and virtual electronic spaces. 

Finally we would like to emphasize that AIDMD, though conceptually simple, 
is still a method in its development stages, with room for algorithmic improvements and new applications. It suffers 
from the usual problems of optimization and overfitting, but with advances in these fields we believe some of these major 
concerns may become less problematic. Some of the issues that need further research are,
\begin{itemize} 
	\item Construction of wavefunction database:
	The AIDMD method relies crucially on the availability of a low energy space of \emph{ab initio} wavefunctions, 
	which despite not being eigenstates, reveal the nature of the effective Hamiltonian. Automating its construction 
	remains somewhat challenging. We propose Slater Jastrow wavefunctions and deforming the orbitals entering the Slater 
	determinant as one way of probing the low energy manifold.
	\item Optimal choice of active orbitals (one body space):
	In the present work, we chose DFT orbitals (in a certain energy window) and localized them to get the one body orbitals. 
	The choice could only be justified based on looking at the trace of the 1-RDM and ensuring that it 
	equalled the expected number of electrons. We expect to use intrinisic atomic orbitals (pointed out to us by G. K. Chan) which 
	mitigate this issue. 
	\item Form of the low energy model Hamiltonian:
	Extensive work in the literature has been devoted to parameterizing compact effective Hamiltonians~\cite{Georges, Oles, Coury}. 
	Since the low energy effective Hamiltonian is not unique, there is not necessarily one right way of downfolding. 
	One could catalogue all functional forms based on previous parameterizations 
	or there may be some merit in determining the minimal description with fewest non zero parameters.
	Note that most models for strong correlation \emph{assume} two body forms. While this appears to work well in practice, 
	it is by not guaranteed on mathematical grounds. (Even though the Coulomb operator is two body, 
	its low energy description need not be.)
	\item Other (non QMC) wavefunction based electronic structure methods:
	There are several wavefunction based quantum chemistry 
	methods for electronic structure like coupled cluster and ab-initio DMRG which work primarily 
	directly in orbital space. These could also be potentially explored in conjunction with AIDMD.
	
\end{itemize} 

In terms of applications, we hope the method will be eventually applied to solids involving transition 
metals and/or rare earth elements, which are in the strongly correlated regime due to the presence of localized orbitals. 
Some challenging areas include,
\begin{itemize} 
	\item Spin models for magnetism: 
	The magnetism associated with strongly correlated Mott insulators, compounded by effects such as geometrical frustration, 
	is highly non trivial and has become an active area of research in its own right. 
	%This is compounded by 
	%the presence of geometric frustration, impurities and defects and 
	%spin-orbital effects. 
	For example, the two dimensional kagome lattice planes in 
	herbertsmithite harbor an exotic topological phase of matter - a "quantum spin liquid". While our understanding is primarily 
	from Heisenberg models~\cite{Yan_Huse_White, Changlani_kagome}, 
	contentious issues plague a complete understanding and connection to experiment (there has been recent progress 
	in downfolding herbertsmithite~\cite{Jeschke}). 
	Other areas of application include pyrochlore iridates and titanates and Kitaev materials. 
	We caution that relevant energy scales associated with magnetism can be $0.1$ meV or less, and no electronic structure 
	method is currently that accurate. It may thus be necessary to perform multiple downfolding steps to map from 
	the Schrodinger equation to a spin model.
	\item Multiband models for superconductors:
	As previously mentioned, unconventional (non BCS) superconductivity in materials such as the cuprates and iron based 
	pntictides has fuelled the study of strongly correlated systems.
	Several parameter sets exist in the literature for multi-band models, but it appears that there is little universal consensus. 
	We are optimistic that additional \emph{ab-initio} inputs from AIDMD will help constrain the part of parameter 
	space relevant for these materials. Trends of the parameters (and the effectiveness of the three band 
	or one band model itself) with pressure dependence and doping remain largely unexplored.  
	%\item Quantum computing (\HJC{tell me if this one is too weird}) : 
	%Quantum computers have been heralded as the future of computing with many promising advances. 
	%There has been progress in the simulation of small molecules in compact basis sets, but scalability (needed for solids) 
	%remains an issue. Downfolding macould help in the reduction in the number of active orbitals 
\end{itemize} 


\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}
